---
title: "phage trial weather data"
author: "Kristi Gdanetz MacCready"
date: '2023-06-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE)
knitr::opts_knit$set(root.dir = "~/Documents/Flower_Microbiome/")
```

# Load libraries 
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
library(weathermetrics)
library(measurements)
library(cowplot)
theme_set(theme_cowplot())
set.seed(125)
```

# Custom functions
### Import solar data
Tables of daily sunrise and sunset times from datetime.org were copied and pasted into tabs of an Excel spreadsheet for location and year combination. Sunrise and Sunset times were converted from AM/PM to 24hr within Excel, and "Year" column was added (format MM-YYYY). The custom function below imports data from a single spreadsheet tab, converts POSIXct (times) to characters, which will be converted to dttm in another function to be use arithmetically. 
```{r}
import_solar <- function(FileLocation, SheetName, Site){
  NewDF <- read_excel(path = FileLocation, sheet = SheetName, col_names = TRUE,
                       col_types = c("date", "numeric", "date", "date")) %>%
  # add missing metadata 
  add_column(Location = Site) %>%
  # format times
  separate(col = "Sunrise", into = c("tmp1", "Sunrise"), sep = " ", remove = TRUE) %>%
  separate(col = "Sunrise", into = c("Rise_hr", "Rise_min", "Rise_Sec"), 
           sep = ":", remove = TRUE) %>%
  separate(col = "Sunset", into = c("tmp2", "Sunset"), sep = " ", remove = TRUE) %>%
  separate(col = "Sunset", into = c("Set_hr", "Set_min", "Set_Sec"), 
           sep = ":", remove = TRUE) %>% 
  separate(col = "Year", into = c("Year", "Month", "tmp3"), sep = "-", remove = TRUE) %>%
  mutate(Month = recode(Month, "05" = "5"),
         across(Year:Day, as.numeric)) %>%
  # drop extra columns 
  select(-tmp1, -tmp2, -tmp3)
  return(NewDF)
}
```

### Import EnviroWeather
CSVs of hourly data for duration of study were downloaded from enviroweather.msu.edu > Data on Demand which transfers to mawn.geo.msu.edu, selected Station > Hourly Data and choose options below to generate a custom report:
  atmp:          Air Temperature [1.5m] (C)
  pcpn:          Precipitation (mm)
  relh:          Relative Humidity [1.5m] (%)
  srad:          Total Solar Flux (kJ/m^2)
  wspd:          Wind Speed [3m] (m/s)
  
Plain text output was copied and pasted into Excel, report and station metadata were removed. Reports from each year and site were pasted as separate tabs in a single spreadsheet. The custom function below imports the spreadsheet containing data downloaded from EnviroWeather. Units are converted to metric, dates and times are converted from POSIXct to characters then numeric, and eventually dttm with another function so times could be used arithmetically. 
```{r}
import_enviro <- function(FileLocation, SheetName, Site){
  NewDF <- read_excel(path = FileLocation, sheet = SheetName, skip = 3,
                    col_names = TRUE, na = c("", " ", "M", "NA"), 
                    col_types = c("date", "date", "numeric", "skip", 
                                  "numeric", "numeric", "numeric", "skip", "skip",
                                  "numeric", "numeric", "numeric", "numeric", "numeric")) %>%
  # rename columns
  rename(Air_temp = atmp,  Precipitation = pcpn,
         Relative_humidity = relh, Wind_speed = wspd) %>% 
  add_column(Location = Site) %>%
  # update metric units 
  mutate(Wind_speed = conv_unit(Wind_speed, from = "m_per_sec", to = "kph"),
         J_m2 = srad*1000 ) %>%
  select(Location, date, time, Air_temp, Precipitation, Relative_humidity, J_m2, Wind_speed) %>%
   # format date & time 
  separate(col = "date", into = c("Year", "Month", "Day"), sep = "-", remove = FALSE) %>%
  separate(col = "time", into = c("tmp1", "Hours"), sep = " ", remove = TRUE) %>%
  separate(col = "Hours", into = c("Hr", "Min", "Sec"), sep = ":", remove = TRUE) %>%
  select(-tmp1) %>%
  mutate(across(Year:Sec, as.numeric))
  return(NewDF)
}
```

### Import NEWA
CSVs of hourly data for the duration of the study were downloaded from newa.cornell.edu and were combined as tabs within a Microsoft Excel spreadsheet. Degree symbols were manually removed in Excel. The custom function below imports the spreadsheet containing data downloaded from NEWA. Units are converted to metric, dates and times are converted from POSIXct to characters to numeric, and eventually dttm with another function so times could be used arithmetically. 
```{r}
import_newa <- function(FileLocation, SheetName, Site) {
  NewDF <- read_excel(path = FileLocation, sheet = SheetName, col_names = TRUE,
                   na = c("", " ", "M", "NA"), 
                   col_types = c("date", "numeric", "numeric", "numeric",
                                 "numeric", "numeric", "numeric", 
                                 "numeric", "numeric", "numeric")) %>%
  # convert degrees, F to C
  mutate(Air_temp = round(conv_unit(`Air Temp F`, from = "F", to = "C"), 
                          digits = 1) ) %>%
  # convert inches to mm 
  mutate(Precipitation = round(conv_unit(`Precipitation (inches)`, from = "inch", to = "mm"), 
                               digits = 0) )  %>% 
  # convert Langleys to Joules / m^2 
  mutate(J_m2 = (`Solar Radiation (langleys)`)*41840) %>%
  # convert mph to kph 
  mutate(Wind_speed = round(conv_unit(`Wind Speed (mph)`, from = "mph", to = "kph"), 
                            digits = 1) ) %>% 
  # add missing metadata 
  add_column(Location = Site) %>% 
  rename(Relative_humidity = "Relative Humidity (%)") %>%
  select(Location, date, everything(), -"Air Temp F", -"Leaf Wetness (minutes)", 
         -"Precipitation (inches)", -"Solar Radiation (langleys)", -"Wind Direction (degrees)", 
         -"Wind Speed (mph)", -"Soil Temp (F)",  -"Soil Tension (kPa)") %>%
  # format date & time 
  separate(col = "date", into = c("Dates", "Hours"), sep = " ", remove = FALSE) %>%
  separate(col = "Dates", into = c("Year", "Month", "Day"), sep = "-", remove = TRUE) %>%
  separate(col = "Hours", into = c("Hr", "Min", "Sec"), sep = ":", remove = TRUE) %>%
  mutate(across(Year:Sec, as.numeric ))
  return(NewDF)
}
```

### Calculate daily and nightly means
(1) Custom function to combine weather and sunrise/sunset data frames. Combines numeric date and time columns to one ddtm column for sunrise and another for sunset times. 
(2 & 3) Custom functions to extract day time and night time measurements from combined data frame. Day is after sunrise and before sunset. Night is before sunrise or after sunset. Calculates daily or nightly means of each measurement. 
```{r}
combine_dfs <- function(NEWAdf, SolDf){
  OutDf <- left_join(NEWAdf, SolDf, by = c("Location", "Year", "Month", "Day")) %>%
    # make datetime variable from individual columns
    mutate(Measure_time = make_datetime(Year, Month, Day, Hr, Min, Sec), 
           Rise_time = make_datetime(Year, Month, Day, Rise_hr, Rise_min, Rise_Sec),
           Set_time = make_datetime(Year, Month, Day, Set_hr, Set_min, Set_Sec) ) %>%
    # drop extra columns 
    select(-date, -Month, -Hr, -Min, -Sec, 
           -Rise_hr, -Rise_min, -Rise_Sec, -Set_hr, -Set_min, -Set_Sec)
  return(OutDf)
  }

calc_day_mean <- function(OutDf){
  # create day time date frame
  DayDf <- OutDf %>%
    # after sunrise and before sunset
    filter(Measure_time > Rise_time & Measure_time < Set_time) %>%
    # calculate day time means 
    group_by(Location, Year, Day) %>%
    summarise(across(where(is.numeric), ~mean(.x, na.rm = TRUE))) %>%
    ungroup() %>%
    mutate(across(where(is.numeric), ~na_if(.x, "NaN")) )
  return(DayDf)
}

calc_nite_mean <- function(OutDf){
  # create night time data frame 
  NiteDf <- OutDf %>%
    # before sunrise or after sunset 
    filter(Measure_time <= Rise_time | Measure_time >= Set_time) %>%
    # calculate night time means
    group_by(Location, Year, Day) %>% 
    summarise(across(where(is.numeric), ~mean(.x, na.rm = TRUE)) )  %>%
    ungroup() %>%
    mutate(across(where(is.numeric), ~na_if(.x, "NaN")) )
  return(NiteDf)
}
```

### Calculate DLI
Daily light integral (DLI) is mols/m^2/day of photosynthetically active radiation (400-700 nm). NEWA collects solar radiation data in Langelys, which was converted to J/m^2 within the import_newa() function. EnviroWeather collects solar radiation data as solar flux (kJ/m^2), kJ was converted to J within the import_enviro() function. The custom function below converts J/m^2 to DLI. Input is the data frame generated by combine_dfs() function. 
```{r}
calc_DLI <- function(OutDf){
  DLIdf <- OutDf %>%
    group_by(Location, Year, Day) %>%
    # Daily total solar flux
    summarise(J_m2_d = sum(J_m2, na.rm = TRUE)) %>% 
    ungroup() %>%
    # 1 W = 1 J/s, divide by the seconds in a day
    mutate(W_m2_s = J_m2_d/86400,
           # 0.51 = 400-700 nm spectrum conversion factor
           umol_m2_s = W_m2_s/0.51, 
           # 86,400 seconds per day
           umol_m2_d = umol_m2_s*86400,
           # 1,000,000 umol per mol
           DLI = umol_m2_d/1000000 ) %>% 
    # replace any missing values (0) with NA
    mutate(across(where(is.numeric), ~na_if(.x, 0)) ) 
  return(DLIdf)
}
```


# New York 
The section below executes each of the custom functions on data from Cornell's Geneva Station. This will generate three data frames of interest, DLI, daily night time averages, daily day time averages. 
### 2018
```{r}
ny18 <- import_newa("weather_data.xlsx", "NY18", "NY")

ny18_sol <- import_solar("Sunset_data_Ea_phage.xlsx", "NY18", "NY") 

ny18_df <- combine_dfs(ny18, ny18_sol)

ny18_DLI <- calc_DLI(ny18_df)

ny18_day <- calc_day_mean(ny18_df)

ny18_nite <- calc_nite_mean(ny18_df)
```

### 2019
```{r}
ny19 <- import_newa("weather_data.xlsx", "NY19", "NY")

ny19_sol <- import_solar("Sunset_data_Ea_phage.xlsx", "NY19", "NY") 

ny19_df <- combine_dfs(ny19, ny19_sol)

ny19_DLI <- calc_DLI(ny19_df)

ny19_day <- calc_day_mean(ny19_df)

ny19_nite <- calc_nite_mean(ny19_df)
```

### 2020
```{r}
ny20 <- import_newa("weather_data.xlsx", "NY20", "NY")

ny20_sol <- import_solar("Sunset_data_Ea_phage.xlsx", "NY20", "NY") 

ny20_df <- combine_dfs(ny20, ny20_sol)

ny20_DLI <- calc_DLI(ny20_df)

ny20_day <- calc_day_mean(ny20_df)

ny20_nite <- calc_nite_mean(ny20_df)
```

### 2021
```{r}
ny21 <- import_newa("weather_data.xlsx", "NY21", "NY")

ny21_sol <- import_solar("Sunset_data_Ea_phage.xlsx", "NY21", "NY") 

ny21_df <- combine_dfs(ny21, ny21_sol)

ny21_DLI <- calc_DLI(ny21_df)

ny21_day <- calc_day_mean(ny21_df)

ny21_nite <- calc_nite_mean(ny21_df)
```


# North Carolina 
The section below executes each of the custom functions on data from NC State Mill's River Research Station. This will generate three data frames of interest, DLI, daily night time averages, daily day time averages. 
### 2019
```{r}
nc19 <- import_newa("weather_data.xlsx", "NC2019", "NC")

nc19_sol <- import_solar("Sunset_data_Ea_phage.xlsx", "NC19", "NC") 

nc19_df <- combine_dfs(nc19, nc19_sol)

nc19_DLI <- calc_DLI(nc19_df)

nc19_day <- calc_day_mean(nc19_df)

nc19_nite <- calc_nite_mean(nc19_df)
```

### 2021
North Carolina 2020 data missing because of COVID. 
```{r}
nc21 <- import_newa("weather_data.xlsx", "NC2021", "NC")

nc21_sol <- import_solar("Sunset_data_Ea_phage.xlsx", "NC21", "NC") 

nc21_df <- combine_dfs(nc21, nc21_sol)

nc21_DLI <- calc_DLI(nc21_df)

nc21_day <- calc_day_mean(nc21_df)

nc21_nite <- calc_nite_mean(nc21_df)
```

###  2022
```{r}
nc22 <- import_newa("weather_data.xlsx", "NC2022", "NC")

nc22_sol <- import_solar("Sunset_data_Ea_phage.xlsx", "NC22", "NC") 

nc22_df <- combine_dfs(nc22, nc22_sol)

nc22_DLI <- calc_DLI(nc22_df)

nc22_day <- calc_day_mean(nc22_df)

nc22_nite <- calc_nite_mean(nc22_df)
```


# Conneticut 
The section below executes each of the custom functions on data from Lockwood Farm CT Agricultural Experiment Station. This will generate three data frames of interest, DLI, daily night time averages, daily day time averages. 
```{r}
ct20 <- import_newa("weather_data.xlsx", "CT2020", "CT")

ct20_sol <- import_solar("Sunset_data_Ea_phage.xlsx", "CT20", "CT") 

ct20_df <- combine_dfs(ct20, ct20_sol)

ct20_DLI <- calc_DLI(ct20_df)

ct20_day <- calc_day_mean(ct20_df)

ct20_nite <- calc_nite_mean(ct20_df)
```


# MSU Hort Farm
The section below executes each of the custom functions on data from MSU Horticulture Farm in East Lansing. This will generate three data frames of interest, DLI, daily night time averages, daily day time averages. 
### 2017
```{r}
msu17 <- import_enviro("weather_data.xlsx", "MSUHort17", "MSU")

msu17_sol <- import_solar("Sunset_data_Ea_phage.xlsx", "EL17", "MSU") 

msu17_df <- combine_dfs(msu17, msu17_sol)

msu17_DLI <- calc_DLI(msu17_df)

msu17_day <- calc_day_mean(msu17_df)

msu17_nite <- calc_nite_mean(msu17_df)
```

### 2018
```{r}
msu18 <- import_enviro("weather_data.xlsx", "MSUHort18", "MSU")

msu18_sol <- import_solar("Sunset_data_Ea_phage.xlsx", "EL18", "MSU") 

msu18_df <- combine_dfs(msu18, msu18_sol)

msu18_DLI <- calc_DLI(msu18_df)

msu18_day <- calc_day_mean(msu18_df)

msu18_nite <- calc_nite_mean(msu18_df)
```

### 2019
```{r}
msu19 <- import_enviro("weather_data.xlsx", "MSUHort19", "MSU")

msu19_sol <- import_solar("Sunset_data_Ea_phage.xlsx", "EL19", "MSU") 

msu19_df <- combine_dfs(msu19, msu19_sol)

msu19_DLI <- calc_DLI(msu19_df)

msu19_day <- calc_day_mean(msu19_df)

msu19_nite <- calc_nite_mean(msu19_df)
```

### 2020
```{r}
msu20 <- import_enviro("weather_data.xlsx", "MSUHort20", "MSU")

msu20_sol <- import_solar("Sunset_data_Ea_phage.xlsx", "EL20", "MSU") 

msu20_df <- combine_dfs(msu20, msu20_sol)

msu20_DLI <- calc_DLI(msu20_df)

msu20_day <- calc_day_mean(msu20_df)

msu20_nite <- calc_nite_mean(msu20_df)
```

### 2021
```{r}
msu21 <- import_enviro("weather_data.xlsx", "MSUHort21", "MSU")

msu21_sol <- import_solar("Sunset_data_Ea_phage.xlsx", "EL21", "MSU") 

msu21_df <- combine_dfs(msu21, msu21_sol)

msu21_DLI <- calc_DLI(msu21_df)

msu21_day <- calc_day_mean(msu21_df)

msu21_nite <- calc_nite_mean(msu21_df)
```


# NWMHRC
The section below executes each of the custom functions on data from Michigan Northwest Horticultre Research Station in Traverse City. This will generate three data frames of interest, DLI, daily night time averages, daily day time averages. 
### 2021
```{r}
tc21 <- import_enviro("weather_data.xlsx", "NWMHRC21", "TC")

tc21_sol <- import_solar("Sunset_data_Ea_phage.xlsx", "TC21", "TC") 

tc21_df <- combine_dfs(tc21, tc21_sol)

tc21_DLI <- calc_DLI(tc21_df)

tc21_day <- calc_day_mean(tc21_df)

tc21_nite <- calc_nite_mean(tc21_df)
```

### 2022
```{r}
tc22 <- import_enviro("weather_data.xlsx", "NWMHRC22", "TC")

tc22_sol <- import_solar("Sunset_data_Ea_phage.xlsx", "TC22", "TC") 

tc22_df <- combine_dfs(tc22, tc22_sol)

tc22_DLI <- calc_DLI(tc22_df)

tc22_day <- calc_day_mean(tc22_df)

tc22_nite <- calc_nite_mean(tc22_df)
```


# Export tables
### DLI
```{r}
# New York
rbind(ny18_DLI, ny19_DLI, ny20_DLI, ny21_DLI) %>%
  select(Location, Year, Day, everything()) %>% 
  mutate(across(4:8, round, digits = 0) ) %>%
  arrange(Year, Day) %>%
  write_delim(., delim = ",", na = "NA", col_names = TRUE, 
              file = "weather_summary_tables/NY_DLI.csv")

# North Carolina
rbind(nc19_DLI, nc21_DLI, nc22_DLI) %>%
  select(Location, Year, Day, everything()) %>% 
  mutate(across(4:8, round, digits = 0) ) %>%
  arrange(Year, Day) %>%
  write_delim(., delim = ",", na = "NA", col_names = TRUE, 
              file = "weather_summary_tables/NC_DLI.csv")

# Conn
ct20_DLI %>%
  select(Location, Year, Day, everything()) %>% 
  mutate(across(4:8, round, digits = 0) ) %>%
  arrange(Year, Day) %>%
  write_delim(., delim = ",", na = "NA", col_names = TRUE, 
              file = "weather_summary_tables/CT_DLI.csv")

# Michigan 
rbind(msu17_DLI, msu18_DLI, msu19_DLI, msu20_DLI, msu21_DLI, tc21_DLI, tc22_DLI) %>%
  select(Location, Year, Day, everything()) %>% 
  mutate(across(4:8, round, digits = 0) ) %>%
  arrange(Year, Day) %>%
  write_delim(., delim = ",", na = "NA", col_names = TRUE, 
              file = "weather_summary_tables/MI_DLI.csv")
```

### Daily & nightly averages
Combine data frames from New York, export as CSV table. 
```{r}
# Day time averages
rbind(ny18_day, ny19_day, ny20_day, ny21_day) %>% 
  select(Location, Year, Day, everything()) %>% 
  mutate(across(4:8, round, digits = 1 ) ) %>%
  arrange(Year, Day) %>%
  write_delim(., delim = ",", na = "NA", col_names = TRUE, 
              file = "weather_summary_tables/NY_day_avg.csv")
# Night time averages
rbind(ny18_nite, ny19_nite, ny20_nite, ny21_nite) %>%
  select(Location, Year, Day, everything()) %>% 
  mutate(across(4:8, round, digits = 1 ) ) %>%
  arrange(Year, Day) %>%
  write_delim(., delim = ",", na = "NA", col_names = TRUE, 
              file = "weather_summary_tables/NY_nite_avg.csv")
```

Combine data frames from North Carolina, export as CSV table. 
```{r}
# Day time averages
rbind(nc19_day, nc21_day, nc22_day) %>%
  select(Location, Year, Day, everything()) %>% 
  mutate(across(4:8, round, digits = 1 ) ) %>%
  arrange(Year, Day) %>%
  write_delim(., delim = ",", na = "NA", col_names = TRUE, 
              file = "weather_summary_tables/NC_day_avg.csv")
# Night time averages
rbind(nc19_nite, nc21_nite, nc22_nite) %>%
  select(Location, Year, Day, everything()) %>% 
  mutate(across(4:8, round, digits = 1 ) ) %>%
  arrange(Year, Day) %>%
  write_delim(., delim = ",", na = "NA", col_names = TRUE, 
              file = "weather_summary_tables/NC_nite_avg.csv")
```

Export tables from CT as CSV table. 
```{r}
ct20_day %>%
  select(Location, Year, Day, everything()) %>% 
  mutate(across(4:8, round, digits = 1 ) ) %>%
  arrange(Year, Day) %>%
  write_delim(., delim = ",", na = "NA", col_names = TRUE, 
              file = "weather_summary_tables/CT_day_avg.csv")

ct20_nite %>%
  select(Location, Year, Day, everything()) %>% 
  mutate(across(4:8, round, digits = 1 ) ) %>%
  arrange(Year, Day) %>%
  write_delim(., delim = ",", na = "NA", col_names = TRUE, 
              file = "weather_summary_tables/CT_nite_avg.csv")
```

Combine data frames from Michigan, export as CSV table. 
```{r}
# Day time averages
rbind(msu17_day, msu18_day, msu19_day, msu20_day, msu21_day, tc21_day, tc22_day) %>%
  select(Location, Year, Day, everything()) %>% 
  mutate(across(4:8, round, digits = 1 ) ) %>%
  arrange(Year, Day) %>%
  write_delim(., delim = ",", na = "NA", col_names = TRUE, 
              file = "weather_summary_tables/MI_day_avg.csv")
# Night time averages
rbind(msu17_nite, msu18_nite, msu19_nite, msu20_nite, msu21_nite, tc21_nite, tc22_nite) %>%
  select(Location, Year, Day, everything()) %>% 
  mutate(across(4:8, round, digits = 1 ) ) %>%
  arrange(Year, Day) %>%
  write_delim(., delim = ",", na = "NA", col_names = TRUE, 
              file = "weather_summary_tables/MI_nite_avg.csv")
```

-----
end
